---
title: "Materials to support the 2023 assessment of black rockfish in waters off Oregon"
output: html_document
---

\

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  fig.pos = "!H",
  out.extra = "",
  collapse = TRUE,
  comment = "#>"
)

library(sa4ss)

dir <- "C:/Users/Jason.Cope/Documents/Github/Sebastes_melanops_OR/docs"
setwd(dir)


```

```{r, results='asis'}
add_figure(
  filein = file.path(dir, "Sebastes_melanops.jpg"), 
  caption = '',
  width = 60,
  height = 60,
  label = 'copper')

``` 


### Join Stakeholder Discussions

The copper rockfish assessment team has organized three informal discussion session during the month of January. The first meeting on Tuesday January 10th from 3 p.m. until 4:30 p.m. will focus on commercial fisheries. The second meeting scheduled Monday January 23rd from 3 p.m. until 4:30 p.m. will focus on recreational fisheries south of Point Conception. The third meeting scheduled Monday January 30th from 3 p.m. until 4:30 p.m. will focus on recreational fisheries south of Point Conception. Please consider joining one of these informal conversations about fisheries for copper rockfish in California.

<!---
[Meeting and Connection Information for January 2023 Discussions](https://github.com/pfmc-assessments/copper_rockfish_2023/blob/main/docs/Copper%20rockfish%20and%20the%20fisheries%20-%20virtual%20discussions.pdf)
-->

<a href="Copper rockfish and the fisheries - virtual discussions.pdf" target="_blank">Meeting and Connection Information for January 2023 Discussions</a>

### Pre-assessment Data Workshop

This year's data workshop will be held virtually. A joint workshop focusing on copper rockfish, black rockfish, and canary rockfish is scheduled for Tuesday January 31 from 9 a.m. until 3 p.m. and Wednesday February 1st from 9 a.m. to 12 p.m. Information to connect to the meeting can be found [online](https://www.pcouncil.org/events/pre-assessment-workshop-for-copper-canary-and-black-rockfish-january-31-amp-february-1-2023/).

[Materials for the pre-assessment data workshop - January 31 - February 1, 2023](data_workshop)

