```{r executive, echo = FALSE}
executive <- list()
executive[["stock"]] <- paste0("This assessment reports the status of ",
  spp, " (_", spp.sci, "_) off the ", coast, 
  " coast using data through xxxx.")
```

# Executive summary{-}

## Stock{-}

`r executive[["stock"]]`
The assessments described in this document apply to the black rockfish (Sebastes melanops) stocks that reside in the waters from Point Conception (34\textdegree27' N latitude) in the south to the U.S. boundary with Canada (approximately 48\textdegree30' N latitude).  Following the consensus recommendations from a preliminary stock assessment workshop in April 2015 (PFMC 2015), the stock assessment team (STAT) decided to prepare separate geographic stock assessments that are spatially stratified with boundaries at the CA/OR border (42\textdegree00' N latitude) and OR/WA border (46\textdegree16' N latitude).  

Black rockfish are also caught from the waters off British Columbia and Alaska, but there have not been any formal assessments of stock status for those areas.

## Catches{-}

Black rockfish are caught by a wide variety of gear types and in recent decades have been a very important target species for recreational charter-boats and private sport anglers in Washington and Oregon, and to a lesser extent in California.  In recent years the recreational fishery has accounted for most of the black rockfish catches (Figure ES-1 to Figure ES-3).  Black rockfish can also be an important component of nearshore commercial fisheries, either as incidental catch by the troll fishery for salmon or as directed catch by jig fisheries for groundfish.  Further, in California and Oregon there are nearshore fisheries that catch and sell fish live for the restaurant trade.  Washington closed nearshore commercial fisheries in state water in late 1990s and never allowed the live-fish fishery to develop. In all states there have been almost no trawl-caught landings of black rockfish in recent years (Table ES-1), but trawl landings in the past were substantial (Figure ES-1 to Figure ES-3).

Detailed reports of commercial landings of black rockfish are generally unavailable prior to 1981, when the Pacific Fishery Information Network (PacFIN) database began.  The catch series prior to 1981 for these assessments were derived by applying available estimates or assumed values for the proportion of black rockfish landings in reported landings of rockfish.  Observer data, which are available only for the past decade, indicate low levels of discarding of black rockfish, generally less than 2% of total catch.

Because of their nearshore distribution and low abundance compared to other rockfish species, black rockfish are unlikely to have ever comprised a large percentage of rockfish landings, but it seems quite certain that they have been more than a trivial component for many years.  Black rockfish were one of only four rockfish species mentioned by scientific name in reports of rockfish landings in Oregon during the 1940s, and they were one of only six rockfish species mentioned by scientific name in reports of rockfish landings in California during the same period. Mentions of black rockfish extend back before the year 1900 in Washington.

## Data and assessment{-}

The last stock assessments for black rockfish were conducted in 2007 for areas north and south of Cape Falcon (45\textdegree46' North latitude). The current assessments assume three areas instead of two, delineated by the state lines as was agreed upon at a pre-assessment and data workshop in March 2015. The prior assessments used Stock Synthesis 2, while the current assessments use Stock Synthesis 
```{r es-modelversion, echo=FALSE} 
if (exists("model")) {
  strsplit(model$SS_version,";")[[1]][1]
}
```
. The Washington base assessment includes a dockside and tag-based CPUE series, but does not include the abundance estimate time series from that same tagging study which was included in the last assessment due to too many violations in the assumptions of abundance estimation. The same two commercial and single recreational fleets are used as in the last assessment for Washington. The Oregon assessment has three commercial fleets and two recreational fleets, while using five surveys and an additional research study for biological compositions. California also has three commercial fleets and 1 recreational fleet with three surveys of abundance, all based on recreational fisheries. All area models include age data as conditional age at lengths. Length compositions are also included in all models.

## Stock biomass and dynamics{-}

RSpawning stock outputs are all at or above limit reference points (Table ES-2. Only California shows declines significantly below this reference point at any point in the time series. California and Washington stocks show a declining population through most of the 20th Century, with stronger declines in the 1980s, and recoveries beginning in the mid-1990s. Oregon stocks follow this pattern, but with a decline in the most recent period. California (33%) is below the target biomass reference point with an increasing biomass trend (Figures ES-4 and ES-5).  The Oregon stock dropped after the quick ramp up of catches in the late 1970s and continued a steady decline until around year 2000, settling in at a stock status around 60% of initial conditions. The Washington stock, currently 43%, dropped below the target biomass by in the early 1980s, then risen above since the late 1990s and has fluctuated above that point through 2014 (Figures ES-8 and ES-9).

## Recruitment{-}

Replace text with
trends and current levels relative to virgin or historic levels and
description of uncertainty.
Include Table for last 10 years.
Include Figure with long-term estimates.

## Exploitation status{-}

Replace text with
total catch divided by exploitable biomass or SPR harvest rate.
Include Table for last 10 years.
Include Figure with trend in f relative to target vs. trend in biomass relative to the target.

## Ecosystem considerations{-}

Replace text with
a summary of reviewed environmental and ecosystem factors that appear to be correlated with stock dynamics.
These may include variability in they physical environment, habitat, competitors, prey, or predators that directly or indirectly affects the stock's status, vital rates (growth, survival, productivity/recruitment) or range and distribution.
Note which, if any, ecosystem factors are used in the assessment and how (e.g., as background information, in data preparations, as data inputs, in decisions about model structure).

## Reference points{-}

Replace text with
management targets and definition of overfishing, including the harvest rate that brings the stock to equilibrium at $B_{40\%}$, i.e., the $B_{MSY}$ proxy and the equilibrium stock size that results from fishing at the default harvest rate, i.e., the $F_{MSY}$ proxy.
Include Table of estimated reference points for ssb, SPR, exploitation rate, and yield based on SSB proxy for MSY, SPR proxy for MSY, and estimated MSY values.

## Management performance{-}

Include Table of most recent 10 years of
catches in comparison with OFL, ABC, HG, and OY/ACL values,
overfishing levels,
actual catch and discard.
Include OFL (encountered), OFL (retained), and OFL (dead) if different due to discard and discard mortality.

## Unresolved problems and major uncertainties{-}

Replace text with
any special issues that complicate scientific assessment, questions about the best model scenario, etc.

## Decision table and projections{-}

Replace text with
projected yields (OFL, ABC, and ACL), spawning biomass, and stock depletion levels for each year.
OFL calculations should be based on the assumption that future catches equal ABCs and not OFLs.

## Scientific uncertainty{-}

Replace text with
the sigma value and the basis for its calculation.

## Research and data needs{-}

Replace text with
information gaps that seriously impede the stock assessment.
