\clearpage

\floatplacement{figure}{H}

# Figures

## Data

<!-- ====================================================================== --> 
<!-- ******************************* Map ********************************** --> 
<!-- ====================================================================== --> 


```{r, results = 'asis'}
add_figure(
filein = file.path(stock_fig_loc, "black_rockfish_map.png"), 
caption = "Map of the management and assessment areas for black rockfish.",
label = 'map',
alt_caption= "Black Rockfish range from California up to Alaska)
```

\newpage

<!-- ====================================================================== --> 
<!-- ******************* Intro section *************************** --> 
<!-- ====================================================================== --> 

<!-- ====================================================================== --> 
<!-- ******************* Data Used in the Model *************************** --> 
<!-- ====================================================================== --> 

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "data_plot.png"), 
caption = "Summary of data sources used in the reference model.  Trawl\\_wdis: Trawl fishery with discards, Non-Trawl\\_wdis: Non Trawl Commercial Fishery with discards",
label = 'data-plot',
alt_caption= "Data can be found in the provided tables and model files")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "catch2 landings stacked.png"), 
caption = "Landings by fleet used in the reference model where catches in metric tons by fleet are stacked. Trawl\\_wdis: Trawl fishery with discards, Non-Trawl\\_wdis: Non Trawl Commercial Fishery with discards",
label = 'catch',
alt_caption= "Removals by fleet can be found in Table 1 and in the model data file")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(index_fig_loc, "qq_NSlog.png"), 
caption = "Q-Q plot for the commercial nearshore logbook index.",
label = 'qq_NSlog',
alt_caption="Raw index data can be found in the stock assessment archive")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(index_fig_loc, "qq_ORBS.png"), 
caption = "Q-Q plot for the recreational ORBS index.",
label = 'qq_ORBS',
alt_caption="Raw index data can be found in the stock assessment archive")
```

\newpage

## Biology

<!-- ====================================================================== -->
<!-- *************************     Biology     **************************** --> 
<!-- ====================================================================== -->

```{r, results = 'asis'}
add_figure(
filein = file.path(bio_fig_loc, "OR_AG_Source_Sex.png"), 
caption = "Observed length-at-age by data source and sex",
label = 'len-age-data-sex',
alt_caption= "The raw data can be found in the stock assessment data archives")
```	

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(bio_fig_loc, "OR_AG_Sex_Year.png"), 
caption = "Observed length-at-age by sex and year. Total samples are indicated in parentheses",
label = 'len-age-sex-year',
alt_caption= "The raw data can be found in the stock assessment data archives")
```	

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(bio_fig_loc, "OR_VBGF_fit.png"), 
caption = "External fits to the observed length-at-age by sex",
label = 'len-age-fit',
alt_caption= "The raw data can be found in the stock assessment data archives")
```	

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(bio_fig_loc, "OR_CV_Sex_plot.png"), 
caption = "Coefficient of variation of length by age by sex. Numbers indicate samples by age and colors indicate sex",
label = 'cv-lt-age',
alt_caption= "The raw data can be found in the stock assessment data archives")
```	

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "bio1_sizeatage.png"), 
caption = "Model estimated length-at-age. Shaded area indicates 95 percent distribution of length-at-age around the estimated growth curve",
label = 'len-age-ss',
alt_caption= "The raw data can be found in the stock assessment data archives")
```


\clearpage


```{r, results = 'asis'}
add_figure(
filein = file.path(bio_fig_loc, "Age1_1plots.png"), 
caption = "Ageing bias plots by reader comparisons",
label = 'age1-plot',
alt_caption= "The raw data can be found in the stock assessment data archives")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(bio_fig_loc, "OR_Reader_Bias_plot.png"), 
caption = "Estimated bias relationships for each considered matrix. Reader 1 is always considered unbiased. Reader 1a and 1b is an intra-reader comparison. B refers to the bias type and S refers to the imprecision type in the model selection for the ageing error matrix",
label = 'age-error-bias',
alt_caption= "The raw data can be found in the stock assessment data archives")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(bio_fig_loc, "OR_Reader_SD_plot.png"), 
caption = "Ageing error matrix standard deviation (SD) values by comparison. B refers to the bias type and S refers to the imprecision type in the model selection for the ageing error matrix",
label = 'age-error-sd',
alt_caption= "The raw data can be found in the stock assessment data archives")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(bio_fig_loc, "Fig1_bio_and_func_color.png"), 
caption = "Biological and functional maturity ogives. Biological maturity ogives are represented by the dashed lines (logistic regression and 95\\% CI is grey, flexible spline and 95\\% CI is light green). Functional maturity ogives are represented by the solid lines (logistic regression and 95\\% CI is grey, flexible spline and 95\\% CI is light blue)",
label = 'bio-fxn-mat-color',
alt_caption= "The data are available upon equest from the authors")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(bio_fig_loc, "Fig2_func_samplesize_color.png"), 
caption = "Functional maturity ogives and sample size. Functional maturity ogives are represented by the solid lines (logistic regression and 95\\% CI is grey, flexible spline and 95\\% CI is light blue) and sample size is denoted by the size of the bubbles)",
label = 'fxn-spline-mat-color',
alt_caption= "The data are available upon equest from the authors")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "bio6_maturity.png"), 
caption = "Maturity as a function of length (cm) used in the reference model",
label = 'maturity',
alt_caption= "The data are available upon equest from the authors")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "bio9_fecundity_len.png"), 
caption = "Fecundity (kg) as a function of length (cm) used in the reference model",
label = 'fecundity',
alt_caption= "The fecundity relationship was provided by E.J. Dick")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(bio_fig_loc, "LW_OR_State_Source_Sex.png"), 
caption = "Sex-specific length (cm)-weight (kg) data for Oregon black rockfish samples by source. MRFSS and ORBS are the ocean boat recreational fishery from early and late periods",
label = 'len-weight-data',
alt_caption= "The raw data are found in the stock assessment archive")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(bio_fig_loc, "LW_lines_States_Sex.png"), 
caption = "Sex-specific length (cm)-weight (kg) estimated power function relationships. Washington state estimate relationships are also provided for comparison",
label = 'len-weight-or-wa',
alt_caption= "The raw data are found in the stock assessment archive")
```

\clearpage

<!-- ====================================================================== -->
<!-- *************************     Indices     **************************** --> 
<!-- ====================================================================== -->

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "index9_standcpueall.png"), 
caption = "The set of standardized indices (each index is scaled to have a mean observation of 1) used in the reference model.",
label = 'sur-summ-all',
alt_caption="Raw index data can be found in the stock assessment archive")
```

\newpage

<!-- ====================================================================== --> 
<!-- ******************* Data Used in the Model *************************** --> 
<!-- ====================================================================== --> 

```{r, results = 'asis'}
add_figure(
filein = file.path(index_fig_loc, "qq_MRHnL.png"), 
caption = "Q-Q plot for the marine reserves hook and line survey.",
label = 'qq_MRHnL',
alt_caption="Raw index data can be found in the stock assessment archive")
```

\newpage


```{r, results = 'asis'}
add_figure(
filein = file.path(index_fig_loc, "index1_cpuedata_SMURF.png"), 
caption = "Index of recruitment from SMURF data collections. Error bars indicate 95\\% uncertainty interval around index values based on the model assumption of lognormal error. This index was not included in the reference model, but was evaluated as a sensitivity.",
label = 'index1_cpuedata_SMURF',
alt_caption="Raw index data can be found in the stock assessment archive")
```

\newpage

## Model Results

<!-- ====================================================================== --> 
<!-- ******************* Bridge Model ************************************* --> 
<!-- ====================================================================== --> 

### Bridging

```{r, results = 'asis'}
add_figure(
  filein = file.path(bridge_fig_loc, "Assess_catch_compare.png"), 
  caption = "Total (top panel), recreational (middle panel), and commercial (bottom panel) removal history used in this assessment compared to the previous (2015) assessment",
  label = 'catch-bridge',
  alt_caption= "These model files are available in the stock assessment archive")

``` 

\newpage

```{r, results = 'asis'}
add_figure(
  filein = file.path(bridge_fig_loc, "OR_SB_comp_plot.png"), 
  caption = "Comparison of spawning output for black rockfish in waters off of Oregon between Stock Synthesis versions 3.24 and 3.30. Uncertainty envelops are 95\\% confidence intervals",
  label = 'ssb-bridge-comps',
  alt_caption= "These model files are available in the stock assessment archive")

``` 

\newpage

```{r, results = 'asis'}
add_figure(
  filein = file.path(bridge_fig_loc, "OR_Dep_comp_plot.png"), 
  caption = "Comparison of relative spawning output for black rockfish in waters off of Oregon between Stock Synthesis versions 3.24 and 3.30. Uncertainty envelops are 95\\% confidence intervals",
  label = 'deps-bridge-comps',
  alt_caption= "These model files are available in the stock assessment archive")

``` 

\clearpage

### Model Development

<!-- ====================================================================== -->
<!-- ***********************  Model Build  ******************************** --> 
<!-- ====================================================================== -->

```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(mod_build_fig_loc, "Mod build stock status/compare4_Bratio_uncertainty.png"),
caption = "Relative spawning output time series by data treatment compared to the reference model. Each scenario is labelled to indicate what data are being used in the scenario. For instance, 'Catch Lengths' is a model with catch and lengths only. 'Dome' referes to a dome-shaped selectivity option for all fisheries. LH refers to life history",
label = 'modbuild-data-depl',
alt_caption= "These model files are available in the stock assessment archive")
```

\newpage


```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(mod_build_fig_loc, "Mod build stock scale/compare2_spawnbio_uncertainty.png"),
caption = "Spawning output (in billions of eggs) time series by data treatment compared to the reference model. Each scenario is labelled to indicate what data are being used in the scenario. For instance, 'Catch Lengths' is a model with catch and lengths only. 'Dome' referes to a dome-shaped selectivity option for all fisheries. LH refers to life history",
label = 'modbuild-data-ssb',
alt_caption= "These model files are available in the stock assessment archive")
```


\newpage

<!-- ============================================================================== -->
<!-- ******************   Pre-STAR  Sensitivities ********************************* --> 
<!-- ============================================================================== -->

```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(sensi_data_fig_loc, "pre-STAR/Sensi_logREplot_SB_Dep_F_MSY.png"),
caption = "Log relative change (log((Model\\_sensi-Model\\_ref)/Model\\_ref)) in data treatment for 5 derived quantities. Colored boxes indicate 95 percent confidence interval of the pre-STAR reference model. See section 3.5.1.1 for more details on each scenario",
label = 'sensi-data-RE_preSTAR'
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(sensi_data_fig_loc, "pre-STAR/compare2_spawnbio_uncertainty.png"),
caption = "Spawning output (in billions of eggs) time series by data treatment compared to the pre-STAR reference model. See section 3.5.1.1 for more details on each scenario",
label = 'sensi-data-ssb_preSTAR'
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(sensi_data_fig_loc, "pre-STAR/compare4_Bratio_uncertainty.png"),
caption = "Relative spawning output time series by data treatment compared to the pre-STAR reference model. See section 3.5.1.1 for more details on each scenario",
label = 'sensi-data-depl_preSTAR'
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(sensi_modspex_fig_loc, "pre-STAR/Sensi_logREplot_SB_Dep_F_MSY.png"),
caption = "Log relative change (log((Model\\_sensi-Model\\_ref)/Model\\_ref)) in model specification scenario for 5 derived quantities. Colored boxes indicate 95 percent confidence interval of the pre-STAR reference model. See section 3.5.1.2 for more details on each scenario",
label = 'sensi-modspec-RE_preSTAR'
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(sensi_modspex_fig_loc, "pre-STAR/compare2_spawnbio_uncertainty.png"),
caption = "Spawning output (in billions of eggs) time series by model specification scenario compared to the pre-STAR reference model. See 'section 3.5.1.2 for more details on each scenario",
label = 'sensi-modspec-ssb_preSTAR'
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(sensi_modspex_fig_loc, "pre-STAR/compare4_Bratio_uncertainty.png"),
caption = "Relative spawning output time series by model specification scenario compared to the pre-STAR reference model. See section 3.5.1.2 for more details on each scenario",
label = 'sensi-modspec-depl_preSTAR'
alt_caption="The model files are available in the stock assessment archive")
```

\newpage



<!-- ============================================================================== -->
<!-- ******************   Pre-STAR Likelihood Profiles **************************** --> 
<!-- ============================================================================== -->

```{r, results = 'asis'}
add_figure(
filein = file.path(likeprof_fig_loc,"pre-STAR/parameter_panel_LnQ_base_Acoustic_Visual(6).png"), 
caption = "Acoustic-visual survey catchability likelihood profile (change in the negative log-likelihood across a range of catchability values) and derived quantities.",
label = 'AVq-profile'
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(likeprof_fig_loc, "pre-STAR/piner_panel_LnQ_base_Acoustic_Visual(6).png"), 
caption = "Acoustic-visual survey catchability likelihood profile for each of the likelihood components",
label = 'AVq-profile-components'
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(likeprof_fig_loc,"pre-STAR/parameter_panel_SR_BH_steep.png"), 
caption = "Beverton-Holt steepness parameter likelihood profile (change in the negative log-likelihood across a range of steepness values) and derived quantities.",
label = 'steepness-profile'
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(likeprof_fig_loc, "pre-STAR/piner_panel_SR_BH_steep.png"), 
caption = "Beverton-Holt steepness parameter likelihood profile for each of the likelihood components",
label = 'steepness-profile-components'
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(likeprof_fig_loc, "pre-STAR/M_fm_multilikelihood_profile.png"), 
caption = "Female and male $M$ multi-parameter likelihood profile and derived quantities. Red lines in the top left figure indicate significantly similar values compared to the reference model. Broken and solid lines in the bottom right figure indicate target and limit referene points, respectively",
label = 'M-multiprofile-combo'
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(likeprof_fig_loc, "pre-STAR/Piner_like_comps.png"), 
caption = "Likelihood component values for the female and male $M$ multi-parameter likelihood profile",
label = 'M-multiprofile-like-components'
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(likeprof_fig_loc, "pre-STAR/LC_Lt.png"), 
caption = "Length likelihood component values for the female and male $M$ multi-parameter likelihood profile",
label = 'M-multiprofile-length-components'
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(likeprof_fig_loc, "pre-STAR/LC_age.png"), 
caption = "Age likelihood component values for the female and male $M$ multi-parameter likelihood profile",
label = 'M-multiprofile-age-components'
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(likeprof_fig_loc, "pre-STAR/LC_survey.png"), 
caption = "Survey likelihood component values for the female and male $M$ multi-parameter likelihood profile",
label = 'M-multiprofile-survey-components'
alt_caption="The model files are available in the stock assessment archive")
```

\newpage


<!-- ====================================================================== -->
<!-- ***********************  Model convergence and Jitters     *********** --> 
<!-- ====================================================================== -->

```{r, results = 'asis'}
add_figure(
filein = file.path(modconv_fig_loc, "jitterplot.png"), 
caption = "Jitter runs for the black rockfish reference model, with jitter run number on the x-axis and -log likelihood value on the y-axis. Blue dot are models that match the likelihood value of the reference model, while red dots deviate from the reference model. All red dots are above the blue dots, indicating no better fit to the reference model was found",
label = 'jitter',
alt_caption= "These model files are available in the stock assessment archive")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(modconv_fig_loc, "pairs_plot_fast.png"), 
caption = "Pairs plots of the fastest mixing parameters from running 2000 posterior draws (and keep every draw) using the random walk Metropolis algorithm. Parameters that show little to no movement are recommended to be fixed to improve model speed and efficiency",
label = 'pairs-plot-fast',
alt_caption= "The model output that produce these plots are available upon request")
```

\clearpage


### Fit to Data

<!-- ====================================================================== -->
<!-- ****************** Fit to the Length Data **************************** --> 
<!-- ====================================================================== -->


```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "comp_lenfit__multi-fleet_comparison.png"), 
caption = "Pearson residuals for each fishing fleet and the MPA survey. Closed bubble are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected)",
label = 'lt-pearson-resids')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "comp_lenfit_data_weighting_TA1.8_Trawl_wdis.png"), 
caption = "Mean length (cm) index from the commercial trawl fishery with 95 percent confidence intervals based on sample sizes and data weighting",
label = 'trawl-mean-len-fit')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "comp_lenfit_data_weighting_TA1.8_Non-Trawl_wdis.png"), 
caption = "Mean length (cm) index from the commercial non-trawl fishery with 95 percent confidence intervals based on sample sizes and data weighting",
label = 'nontrawl-mean-len-fit')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "comp_lenfit_data_weighting_TA1.8_Ocean.png"), 
caption = "Mean length (cm) index from the recreational ocean boat fishery with 95 percent confidence intervals based on sample sizes and data weighting",
label = 'ocean-mean-len-fit')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "comp_lenfit_data_weighting_TA1.8_Shore.png"), 
caption = "Mean length (cm) index from the recreational shore-based fishery with 95 percent confidence intervals based on sample sizes and data weighting",
label = 'shore-mean-len-fit')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "comp_lenfit_data_weighting_TA1.8_MPA.png"), 
caption = "Mean length (cm) index from the MPA survey with 95 percent confidence intervals based on sample sizes and data weighting",
label = 'mpa-mean-len-fit')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "comp_lenfit__aggregated_across_time.png"), 
caption = "Aggregated length (cm) compositions over all years",
label = 'agg-len-fit')
```

\newpage

<!-- ====================================================================== -->
<!-- ****************** Fit to the Age Data ******************************* --> 
<!-- ====================================================================== -->

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "comp_condAALfit_data_weighting_TA1.8_condAgeTrawl_wdis.png"), 
caption = "Mean age from conditional age-at-length data for the commercial trawl fishery",
label = 'trawl-mean-caal')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "comp_condAALfit_data_weighting_TA1.8_condAgeNon-Trawl_wdis.png"), 
caption = "Mean age observations from the conditional age-at-length data from the non-trawl commercial fishery",
label = 'nontrawl-mean-caal')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "comp_condAALfit_data_weighting_TA1.8_condAgeOcean.png"), 
caption = "Mean age observations from the conditional age-at-length data from the ocean boat fishery",
label = 'ocean-mean-caal')
```

\newpage

<!-- ====================================================================== -->
<!-- ****************** Fit to the Index*********************************** --> 
<!-- ====================================================================== -->


```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "index2_cpuefit_Non-Trawl_wdis.png"),
caption = "Fit to the non-trawl commercial survey index of abundance",
label = 'nontrawl-index-fit')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "index2_cpuefit_MRFSS.png"),
caption = "Fit to the MRFSS recreational survey index of abundance",
label = 'mrfss-index-fit')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "index2_cpuefit_Ocean.png"),
caption = "Fit to the ORBS recreational survey index of abundance",
label = 'orbs-index-fit')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "index2_cpuefit_Tag.png"),
caption = "Fit to the tagging survey index of abundance",
label = 'tag-index-fit')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "index2_cpuefit_MPA.png"),
caption = "Fit to the MPA survey index of abundance",
label = 'mpa-index-fit')
```

\newpage


### Estimated Recruitment

<!-- ====================================================================== -->
<!-- ******************     Recruitment    ***************************** --> 
<!-- ====================================================================== -->

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "ts11_Age-0_recruits_(1000s)_with_95_asymptotic_intervals.png"), 
caption = "Estimated time series of age-0 recruits (1000s)",
label = 'recruits')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "recdevs2_withbars.png"), 
caption = "Estimated time series of recruitment deviations",
label = 'rec-devs')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "recdevs3_varcheck.png"), 
caption = "Recruitment deviations variance by year. This plot tracks the information content contained in each recruitment deviation. Values below the red line (assumed recruitment variability) indicates years with more informed recruitment deviations",
label = 'rec-devs-sigmas')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "SR_curve.png"), 
caption = "Stock-recruit curve. Point colors indicate year, with warmer colors indicating earlier years and cooler colors in showing later years",
label = 'bh-curve')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "recruit_fit_bias_adjust.png"),
caption = "Recruitment bias adjustment applied in the reference model",
label = 'bias-adj')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "recdevs3_varcheck.png"),
caption = "Recruitment deviations variance check. Low standard deviations indicate years with informative deviations ",
label = 'varcheck')
```

\newpage

### Selectivity

<!-- ====================================================================== -->
<!-- ******************     Selectivity    ***************************** --> 
<!-- ====================================================================== -->

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "sel01_multiple_fleets_length1.png"),
caption = "Length-based selectivity curves for each fleet and survey",
label = 'fleet-selectivity')
```

\newpage

### Time Series

<!-- ======================================================= -->  
<!-- ****************** Time Series ************************ --> 
<!-- ======================================================= -->


```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "ts7_Spawning_output_with_95_asymptotic_intervals_intervals.png"), 
caption = "Estimated time series of spawning output (in billions of eggs)",
label = 'ssb')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "ts1_Total_biomass_(mt).png"), 
caption = "Estimated time series of total biomass (mt)",
label = 'tot-bio')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "ts9_Relative_spawning_output_intervals.png"), 
caption = "Estimated time series of fraction of unfished spawning output",
label = 'depl')
```

\newpage

### Comparison to Previous Assessments

```{r, results = 'asis'}
add_figure(
filein = file.path(comp_ass_loc, "compare2_spawnbio_uncertainty.png"), 
caption = "Comparison of the time series of spawning output between the 2015 and 2023 assessment results",
label = 'comp-ssb')
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(comp_ass_loc, "compare4_Bratio_uncertainty.png"), 
caption = "Comparison of the time series of relative spawning output between the 2015 and 2023 assessment results",
label = 'comp-depl')
```

### Sensitivities

<!-- ====================================================================== -->
<!-- ******************       Sensitivity     ***************************** --> 
<!-- ====================================================================== -->

```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(sensi_data_fig_loc, "Sensi_logREplot_SB_Dep_F_Yield.png"),
caption = "Log relative change (log((Model\\_sensi-Model\\_ref)/Model\\_ref)) in data treatment for 5 derived quantities. Colored boxes indicate 95 percent confidence interval of the reference model. See 'Sensitivity Analysis' section for more details on each scenario",
label = 'sensi-data-RE',
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(sensi_data_fig_loc, "compare2_spawnbio_uncertainty.png"),
caption = "Spawning output (in billions of eggs) time series by data treatment compared to the reference model. See 'Sensitivity Analysis' section for more details on each scenario",
label = 'sensi-data-ssb',
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(sensi_data_fig_loc, "compare4_Bratio_uncertainty.png"),
caption = "Relative spawning output time series by data treatment compared to the reference model. See 'Sensitivity Analysis' section for more details on each scenario",
label = 'sensi-data-depl',
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(sexratio_fig_loc, "sex3.png"),
caption = "Sex ratio by age for the model using the sex=3 option",
label = 'sexratioSex3',
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(sexratio_fig_loc, "sex12.png"),
caption = "Sex ratio by age for the reference model (sex option 1 and 2)",
label = 'sexratioSex12',
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(sexratio_fig_loc, "sex12_logsel.png"),
caption = "Sex ratio by age for the model using sex = 1 and 2, and logistic selectivity",
label = 'sexratio12Log',
alt_caption="The model files are available in the stock assessment archive")
```

\newpage


```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(sensi_modspex_fig_loc, "Sensi_logREplot_SB_Dep_F_Yield.png"),
caption = "Log relative change (log((Model\\_sensi-Model\\_ref)/Model\\_ref)) in model specification scenario for 5 derived quantities. Colored boxes indicate 95 percent confidence interval of the reference model. See 'Sensitivity Analysis' section for more details on each scenario",
label = 'sensi-modspec-RE',
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(sensi_modspex_fig_loc, "compare2_spawnbio_uncertainty.png"),
caption = "Spawning output (in billions of eggs) time series by model specification scenario compared to the reference model. See 'Sensitivity Analysis' section for more details on each scenario",
label = 'sensi-modspec-ssb',
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, echo=FALSE, results='asis'}
add_figure(
filein = file.path(sensi_modspex_fig_loc, "compare4_Bratio_uncertainty.png"),
caption = "Relative spawning output time series by model specification scenario compared to the reference model. See 'Sensitivity Analysis' section for more details on each scenario",
label = 'sensi-modspec-depl',
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

### Likelihood Profiles

<!-- ============================================================================== -->
<!-- ******************      Likelihood Profiles      ***************************** --> 
<!-- ============================================================================== -->

```{r, results = 'asis'}
add_figure(
filein = file.path(likeprof_fig_loc,"parameter_panel_SR_LN(R0).png"), 
caption = "Initial recruitment ($lnR_0$) likelihood profile (change in the negative log-likelihood across a range of $ln(R0)$ values) and derived quantities. Red line in the top left figure indicates the significance level in likelihood difference",
label = 'R0-profile',
alt_caption="The model files are available in the stock assessment archive")
```

\pagebreak

```{r, results = 'asis'}
add_figure(
filein = file.path(likeprof_fig_loc, "piner_panel_SR_LN(R0).png"), 
caption = "Initial recruitment ($ln(R0)$) likelihood profile for each of the likelihood components",
label = 'R0-profile-components',
alt_caption="The model files are available in the stock assessment archive")
```

\pagebreak

```{r, results = 'asis'}
add_figure(
filein = file.path(likeprof_fig_loc,"parameter_panel_SR_BH_steep.png"), 
caption = "Beverton-Holt steepness parameter likelihood profile (change in the negative log-likelihood across a range of steepness values) and derived quantities. Red line in the top left figure indicates the significance level in likelihood difference",
label = 'steepness-profile',
alt_caption="The model files are available in the stock assessment archive")
```

\pagebreak

```{r, results = 'asis'}
add_figure(
filein = file.path(likeprof_fig_loc, "piner_panel_SR_BH_steep.png"), 
caption = "Beverton-Holt steepness parameter likelihood profile for each of the likelihood components",
label = 'steepness-profile-components',
alt_caption="The model files are available in the stock assessment archive")
```

\pagebreak

```{r, results = 'asis'}
add_figure(
filein = file.path(likeprof_fig_loc, "M_fm_multilikelihood_profile.png"), 
caption = "Female and male $M$ multi-parameter likelihood profile and derived quantities. Red lines in the top left figure indicate significantly similar values compared to the reference model. Broken and solid lines in the bottom right figure indicate target and limit referene points, respectively",
label = 'M-multiprofile-combo',
alt_caption="The model files are available in the stock assessment archive")
```

\pagebreak

```{r, results = 'asis'}
add_figure(
filein = file.path(likeprof_fig_loc,"LC_plots.png"), 
caption = "Likelihood values by component and within components for the female and male $M$ multi-parameter likelihood profile",
label = 'M-multiprofile-like-components',
alt_caption="The model files are available in the stock assessment archive")
```

\newpage


### Retrospectives

<!-- ====================================================================== -->
<!-- ******************     Retrospectives    ***************************** --> 
<!-- ====================================================================== -->

```{r, results = 'asis'}
add_figure(
filein = file.path(retro_loc,  "compare2_spawnbio_uncertainty.png"), 
caption = "Change in the estimate of spawning output when the most recent 5 years of data area removed sequentially",
label = 'retro-ssb',
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(retro_loc,  "compare4_Bratio_uncertainty.png"),
caption = "Change in the estimate of fraction unfished when the most recent 5 years of data area removed sequentially",
label = 'retro-depl',
alt_caption="The model files are available in the stock assessment archive")
```

\newpage


```{r, results = 'asis'}
add_figure(
filein = file.path(retro_loc,  "retro_percent_difference_3_panel.png"),
caption = "Relative error from the reference model for each of the 5 data peels in the restrospective analysis for 3 derived outputs",
label = 'retro-RE_comps',
alt_caption="The model files are available in the stock assessment archive")
```

\newpage

### Reference Points and Forecasts

<!-- ====================================================================== -->
<!-- ******************    Reference Points    **************************** --> 
<!-- ====================================================================== -->


```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "SPR2_minusSPRseries.png"), 
caption = "Estimated 1 - relative spawning ratio (SPR) by year",
label = '1-spr',
alt_caption="Values behind this figure are found in the model output Report file")
```

\clearpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "SPR4_phase.png"), 
caption = "Phase plot of the relative biomass (also referred to as fraction unfished) versus the SPR ratio where each point represents the biomass ratio at the start of the year and the relative fishing intensity in that same year. Lines through the final point show the 95 percent intervals based on the asymptotic uncertainty for each dimension. The shaded ellipse is a 95 percent region which accounts for the estimated correlations between the biomass ratio and SPR ratio",
label = 'phase',
alt_caption="Values behind this figure are found in the model output Report file")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "yield2_yield_curve_with_refpoints.png"), 
caption = "Equilibrium yield curve for the reference model. Values are based on the 2020
fishery selectivities and with steepness fixed at 0.72",
label = 'yield',
alt_caption="Values behind this figure are found in the model output Report file")
```

\newpage

```{r, results = 'asis'}
add_figure(
filein = file.path(manage_loc, "compare2_spawnbio_uncertainty.png"), 
caption = "Time series of spawning output (with 95\\% uncertainty envelops) for three states of nature, based on the treatment of the acoustic visual survey catchability, of Oregon black rockfish",
label = 'stateofnature_SO',
alt_caption="Values behind this figure can be found in the Decision Table")
```

```{r, results = 'asis'}
add_figure(
filein = file.path(manage_loc, "compare4_Bratio_uncertainty.png"), 
caption = "Time series of relative stock status (with 95\\% uncertainty envelops) for three states of nature, based on the treatment of the acoustic visual survey catchability, of Oregon black rockfish",
label = 'stateofnature_depl',
alt_caption="Values behind this figure can be found in the Decision Table")
```

\newpage
