```{r executive, echo = FALSE}
executive <- list()
executive[["stock"]] <- paste0("This assessment reports the status of ",
  spp, " (_", spp.sci, "_) off the ", coast, 
  " coast using data through xxxx.")
```

# Executive summary{-}

## Stock{-}

This assessment reports the status of `r spp` (*Sebastes melanops*) off Oregon state using data through `r endyr`. `r Spp` are also found in California and Washington waters of the U.S. West Coast, and those are treated in separate area-based stock assessments given different management considerations and exploitation histories as discussed at the preliminary stock assessment workshop in February 2023 [@PFMC_dataworkshop_2023]. There is substantial biogeographic separation in the populations off Oregon and Washington, thus justifying separation of those populations into different management units and stock assessments. `r spp` are also caught from the waters off British Columbia and Alaska, but there have not been any formal assessments of stock status for those areas in recent years. Genetic studies of stock structure indicate fish in Alaska are more differentiated than those along the contiguous west coast of the United States, and that genetic diversity varies in a non-systematic way from California to Oregon [@hess_comparative_2023].

## Removals{-}

`r spp` are caught by a wide variety of gear types in Oregon and in recent decades have been a very important target species for recreational charter-boats and private sport anglers (i.e., ocean boat fleet).  In recent years the recreational fishery has accounted for most of the `r spp` catches (Figure \ref{fig:es-catch}).  `r spp` are also be an important component of nearshore commercial fisheries, either as incidental catch by the troll fishery for salmon or as directed catch by fixed gear fisheries for groundfish.  Further, there are nearshore commercial fisheries that catch and sell fish live for the restaurant trade. 

Historical catch reconstructions for Oregon show that catch of `r spp` was limited until the development of trawl fisheries around World War II in the 1940s. Trawl landings were consistent through the 1970s and landings from other gear types were minimal (Figure \ref{fig:es-catch}). There has been almost no trawl-caught landings of `r spp` in recent years (Table \ref{tab:removalsES}). Modern-era, detailed catch estimates are available from Pacific Fishery Information Network (PacFIN) database for commercial landings (1987 - 2022) and from the Recreational Fishery Information Network (RecFIN) (2001 - 2022). Historical recreational catch estimates were provided by ODFW (1979 - 2000). Generally, the historical catch series were derived by applying available estimates or assumed values for the proportion of `r spp` landings in reported landings of rockfish.  Observer data, which are available since the early 2000s, indicate low levels of discarding of `r spp`, generally less than 2% of total catch. While `r spp` are unlikely to have ever comprised a large percentage of overall rockfish landings due to their low abundance compared to other rockfish species, it seems plausible that they have been more than a trivial component due to their nearshore distribution for many years.  `r spp` were one of only four rockfish species mentioned by scientific name in reports of rockfish landings in Oregon during the 1940s.

Overall, removals of `r spp` remained relatively low (generally less than 100 mt) until the mid to late 1970s when landings quickly increased (approximately eight-fold, on average) with the expansion of the recreational fishery. Since the 1980s, removals have consistently fluctuated between 300 and 600 mt (no major trend), comprising mostly of removals from the ocean boat recreational fleet and the non-trawl commercial fleet (Figure \ref{fig:es-catch}).  


\input{tex_tables/a_Catches_ES.tex}


```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "catch2 landings stacked.png"), 
caption = "Landings by fleet used in the reference model where catches in metric tons by fleet are stacked. wdis- With Discards",
label = 'es-catch')
```

\clearpage

## Data and Assessment{-}

The first `r spp` stock assessment along the west coast of the United States that included the majority of Oregon waters was completed in 1993, covering the area south of Cape Falcon, Oregon to north of Point Piedros Blancos, California [@sampson_status_2007]. Subsequent assessments that included Oregon waters were done in 2003 and 2007. In 2015, the last assessment prior to the current one was completed that included Oregon waters only as one of three (also Washington and California) separate assessment areas delineated by state lines [@cope_assessments_2016]. Similarly, this assessment treats Oregon waters as a single assessment area. The previous two assessments used Stock Synthesis software, as does this one (version 3.30.21.00).

This assessment integrates data and information from multiple sources into one modeling framework. The stock assessment model for `r spp` is informed by catch data from two commercial fleets and two recreational fleets, six abundance indices, five sets of length composition data, and three sets of conditional age-at-length compositions. It also uses multiple ageing error matrices to incorporate ageing imprecision and applies fixed parameterizations of weight-at-length, maturity-at-length, fecundity-at-length, the Beverton-Holt stock-recruitment steepness value, and recruitment variability. Life history parameters were sex-specific (i.e., a two-sex model) with natural mortality fixed at external estimates and growth and recruitment parameters estimated.  Additional parameters that were estimated include initial population scale ($lnR_0$), selectivity for each fishery and survey, and extra survey variance. Selectivity paramters were estimated using the length composition data and then were fixed in the final model that used marginal age compostion data (instead of lengths) to lessen the effect lengths were having on the assessment. The reference model was tuned to account for the weighting of the length and age data and index variances (which was estimated), as well as the specification of the recruitment bias adjustments.  Derived quantities include, among other things, the time series of spawning output, age and size structure, and current and projected future stock status. The model covers the years `r startyr` to `r endyr`, with a 12 year forecast beginning in `r endyr + 1`.

Within model uncertainty is explicitly included in this assessment by parameter estimation uncertainty, while among model uncertainty is explored through sensitivity analyses addressing alternative input assumptions such as data treatment and weighting, and model specification sensitivity to the treatment of life history parameters, selectivity, recruitment, and survey catchability. A reference model was selected that best fit the observed data while concomitantly balancing the desire to capture the central tendency across those sources of uncertainty, ensure model realism and tractability, and promote robustness to potential model misspecification.

## Stock Biomass and Dynamics{-}

Spawning output (in millions of eggs; meggs) instead of spawning biomass is used to report the functionally mature population scale because fecundity is nonlinearly related to body female weight. The estimated spawning output at the beginning of 2023 was `r ssb_curr` meggs (~95 percent asymptotic intervals: `r ssb_curr_low` to `r ssb_curr_hi` meggs, Table \ref{tab:ssbES} and Figure \ref{fig:es-ssb}), which when compared to unfished spawning output (`r ssb0` meggs) gives a relative stock status level of `r round(100*depl_curr,0)` percent (~95 percent asymptotic intervals: `r round(100*depl_curr_low,0)` to `r round(100*depl_curr_hi,0)` percent, Figure \ref{fig:es-depl}).  Overall, spawning output declined with the onset of increasing commercial removals in the 1960s and continued to decline with the increase in recreational catches through the 1990s, even dropping below the target relative stock size from 1992 to 2014, before steadily increasing back above target since that time. The largest of the estimated recruitment pulses occurred in 2008 and was followed by several above average recruitment years in the early 2010s, which contributed to the increase in spawning output through the mid to late 2010s. The minimum relative stock size of `r round(100*as.numeric(depl_min),0)` percent of unfished levels is estimated to have occurred in `r depl_min_yr`. Accordingly, the stock has not been below the minimum stock size threshold (i.e., never overfished based on median estimates). Currently the stock is estimated above the management target of $SO_{40\%}$ in 2023 and is estimated to have remained above the target since 2014 (Table \ref{tab:ssbES} and Figure \ref{fig:es-depl}).

\input{tex_tables/b_SSB_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots",
"ts7_Spawning_output_with_95_asymptotic_intervals_intervals.png"), 
caption = "Estimated time series of spawning output (circles and line: median; light broken lines: 95 percent intervals) for the reference model",
label = 'es-ssb')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "ts9_Relative_spawning_output_intervals.png"), 
caption = "Estimated time series of fraction of unfished spawning output (circles and line: median; light broken lines: 95 percent intervals) for the reference model",
label = 'es-depl')
```

\clearpage

## Recruitment{-}

Recruitment is informed by the data and estimated from 1980 to 2017, before and after which it is assumed taken from the stock-recruit relationship (Table \ref{tab:recrES} and Figure \ref{fig:es-recruits}).  The highest recruitment years occurred in 1999, 2000, 2008, 2010, 2013, and 2016. The large 2008 and 2016 year classes, as well as several above average year classes in the early 2010s, contributed to the recent increase in `r spp` biomass. Recruitment is informed by composition data and six relative abundance indices.  The 2015 stock assessment did not estimate deviations from the stock-recruitment curve. While the `r spp` stock has been reduced to levels that theoretically would provide some information on how recruitment compensation changes across spawning output levels (i.e., inform the steepness parameter), the assessment model could not adequately estimate a reasonable steepness parameter.  Thus, recruitment is based on a fixed assumption about steepness ($h$ = 0.72) and recruitment variability ($\sigma_R$ = 0.6).

\input{tex_tables/c_Recr_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "ts11_Age-0_recruits_(1000s)_with_95_asymptotic_intervals.png"), 
caption = "Estimated time series of age-0 recruits (1000s) for the reference model with 95 percent intervals",
label = 'es-recruits')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "recdevs2_withbars.png"), 
caption = "Estimated time series of recruitment deviations",
label = 'es-recdev')
```

## Exploitation Status{-}

Fishing intensity, as measured by the SPR rate (1 - SPR), has fluctuated around the estimated SPR rate fishing intensity target of 0.50 (1 - $\text{SPR}_{50\%}$) since 1980. Fishing was consistently at or above the target rate from 1988 to 2006 and has been slightly above it over the past 5 years as well (Table \ref{tab:exploitES} and Figures \ref{fig:es-1-spr} and \ref{fig:es-phase}). The steepness value of 0.72 indicates that a lower value of SPR (or equivalently a higher fishing intensity than $\text{SPR}_{50\%}$) would be consistent with the spawning output-based target of ($\text{SO}_{40\%}$) for sustainable removals. Trends in fishing intensity largely mirrored that of landings until the 1990s, after which recruitment pulses countered the catches somewhat to lower overall fishing intensity (though mostly still at or above target; Figure \ref{fig:es-1-spr}). The maximum fishing intensity (1 - SPR) was `r spr_max` in `r spr_max_yr`, which is well above the target of 0.50. The current level of 0.59 for 2022 is above that target. Fishing intensity over the past decade has ranged between 0.50 and 0.62 and the exploitation rate (range of 0.07 - 0.09, Table \ref{tab:exploitES}) has come down since the time series high of 0.12 in 1992. Current estimates indicate that `r spp` spawning output is greater than than the target biomass level ($\text{SO}_{40\%}$), though fishing intensity remains above the target $F_{MSY}$ proxy harvest rate of 1 - $\text{SPR}_{50\%}$ (Figure \ref{fig:es-phase}).

\input{tex_tables/d_SPR_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "SPR2_minusSPRseries.png"), 
caption = "Estimated 1 - relative spawning ratio (SPR) by year for the reference model. The management target is plotted as a red horizontal line and values above this reflect harvest in excess of the proxy harvest rate",
label = 'es-1-spr')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "SPR4_phase.png"), 
caption = "Phase plot of estimated 1-SPR versus fraction unfished for the reference model",
label = 'es-phase')
```

\clearpage


## Ecosystem Considerations{-}

This stock assessment does not explicitly incorporate trophic interactions, habitat factors or environmental factors into the assessment model. More predation, diet and habitat work, and mechanistic linkages to environmental conditions would be needed to incorporate these elements into the stock assessment and should remain a priority. McClure et al. [-@mcclure_vulnerability_2023] report the climate vulnerability for several west coast groundfishes, including `r spp`. `r spp` demonstrated both high biological sensitivity and high climate exposure risk, to give it an overall high vulnerability score to climate change. This result should also be considered with the fact that, like many rockfishes, periods of low productivity are not unusual to `r spp` and their extended longevity (though admittedly this seems shorter than previously believed and should be reconsidered) has historically allowed them to wait for advantageous productivity periods. Additional stressors such as fishing and climate change that possibly truncate longevity could bring significant challenges to population sustainability.

## Reference Points{-}

Reference points were based on the rockfish FMSY proxy ($\text{SPR}_{50\%}$), target relative spawning output (40%), and estimated selectivity and catch for each fleet (Table \ref{tab:referenceES}). The yield values are lower than the previous assessment for similar reference points due to updated life history estimates. The proxy FMSY values of management quantities, based on exploitation rates that achieve $\text{SPR}_{50\%}$ and $\text{SO}_{40\%}$, are more conservative (lower) than the exploitation rate at estimated MSY. Sustainable total yield using the proxy $\text{SPR}_{50\%}$ is `r format(model$derived_quants[model$derived_quants$Label == 'Dead_Catch_SPR', 'Value'], digits = 3, scientific = FALSE, big.mark = ',')` mt. The spawning output equivalent to 40 percent of the unfished spawning output ($\text{SO}_{40\%}$) was `r format(model$derived_quants[model$derived_quants$Label == 'SSB_Btgt', 'Value'], digits = 4, scientific = FALSE, big.mark = ',')` meggs. 

Recent removals have been above the point estimate of potential long-term yields calculated using an $\text{SPR}_{50\%}$ reference point, though the population size has continued to increase over recent years due to several above average recruitments. The equilibrium estimates of yield relative to biomass based on a steepness value fixed at `r round(model$parameters[model$parameters$Label == "SR_BH_steep","Value"], 2)` are provided in Figure \ref{fig:es-yield}, where vertical dashed lines indicate the estimate of fraction unfished at the start of `r endyr +1` (current) and the estimated management targets calculated based on the relative target biomass (B target), the SPR target, and the maximum sustainable yield (MSY). 

The 2023 spawning output relative to unfished equilibrium spawning output is above (43 percent) the management target of 40 percent of unfished spawning output (i.e., 1.08 times above the target spawning output; Figure \ref{fig:es-phase}). The relative biomass and the ratio of the estimated SPR to the management target ($\text{SPR}_{50\%}$) across all model years are shown in Figure \ref{fig:es-phase} where warmer colors (red) represent early years and colder colors (blue) represent recent years.  There have been periods where the stock status has decreased below the target and fishing intensity has been higher than the target fishing intensity based on $\text{SPR}_{50\%}$. Fishing intensity during the most recent year (2022) is estimated to be above the fishing intensity target (1.18 times $\text{SPR}_{50\%}$; Figure \ref{fig:es-phase}).
 

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "yield2_yield_curve_with_refpoints.png"), 
caption = "Equilibrium yield curve for the reference model. Values are based on (the time invariant)
fishery selectivities and with steepness fixed at 0.72",
label = 'es-yield')
```

\clearpage

\input{tex_tables/e_ReferencePoints_ES.tex}

\clearpage

## Management Performance{-}

`r spp` removals have generally been below the equivalent ABC-ACL over the recent decade, with the exception of 2017 and 2022 when removals were slightly higher (Table \ref{tab:manageES}). Exploitation on `r spp` increased starting around 1940 and reached a high in the late 1970s. Since that time, catch has mostly fluctuated between 300 and 500 mt per year, with some years exceeding 600 mt. Removals have averaged 483 mt over the past decade. The last ten years of `r spp` acceptable biological catch (ABC) and annual catch limit (ACL) have been set, by definition, below the overfishing limit (OFL) (Table \ref{tab:manageES}). Prior to 2017, management specifications were set for `r spp` in California and Oregon waters combined.  The `r spp` OFL has not been exceeded by Oregon removals over the past decade. The ACL and ABC were exceeded in 2017 and 2022 by a small margin, but were more than offset by underages in intervening years.

\input{tex_tables/f_Manage_ES.tex}


## Unresolved Problems and Major Uncertainties{-}

The biggest uncertainty and unresolved conflict is trying to reconcile the signal in the biological data (which suggests a lower population size and status) versus the acoustic and tag surveys (which suggest high stock sizes and status). This is the major issue the current assessment is confronting. Another acoustic-visual survey data point could help resolve how much uncertainty there is in the estimate. The lack of contrast in the biological data, despite large sample sizes, is another barrier to interpreting the current conditions, though given models using only biological data, the signal seems clear that the population could be at a lower stock status.

## Scientific Uncertainty{-}

The biggest uncertainty and unresolved conflict is trying to reconcile the signal in the biological data (which want a lower population size and status) versus the acoustic and tag surveys (which want high stock sizes and status). This is the major issue the current assessment is confronting. Another acoustic-visual survey data point could help resolve how much uncertainty there is in the estimate. The lack of contrast in the length composition data, despite large sample sizes, is another barrier to interpreting the current conditions, as lengths drive the model to a lower stock status, while the ages push it to higher stock status, but not as high as the acoustic-visual survey would have it.

## Harvest Projections and Decision Table{-}

The `r spp` assessment is being considered as a category 1 assessment with a $P^*$ = 0.45, sigma = 0.50, and a time-varying buffer applied to the OFL.  These multipliers are also combined with the rockfish MSY proxy of FSPR=50% MSY and the 40-10 harvest control rule to calculate OFLs, ABCs and ACLs. A twelve year (2023-2034) projection of the reference model using these specifications along with input removals for 2023 and 2024 provided by the Groundfish Management Team is provided in Table \ref{tab:projectionES}. 

\input{tex_tables/g_Projections_ES.tex}
 
Uncertainty in management quantities for the reference model was characterized by exploring various model specifications in a decision table. Initial explorations considered alternative specifications of catchability for the acoustic-visual survey and well as weighting of the biological data. The final states of nature were defined as follows: the high state of nature was the pre-STAR reference model that assume the q on the acoustic-visual survey is 1.82, the highest value deemed reasonable by the acoustic survey team; the low state of nature assumes q = 1 for the acoustic-visual survey, but allows the large CV (45%) on the index, thus allowing the length composition data to drive the model to a lower scale and stock status. The catch rows assume P* values of 0.45 and 0.4, then a constant catch using the yield at FSPR=0.5. The resultant decision table is provided in Table \ref{tab:es-dec-tab}.

\clearpage

\input{tex_tables/decision_table_es.tex}

\clearpage


## Research and Data Needs{-}

Recommended avenues for research to help improve future `r spp` stock assessments are listed below (in no particular order). Responses to research and data needs identified in the last assessment can be found in Section 4.7. 

1. Continue work on the investigation into the movement, and behavior or mortality of older (> age 10) females to further reconcile their absence in fisheries data. 
2. Conduct population genetics studies on fish observed off of the continental shelf (middle of the gyre and at sea mounts) to determine their association with the nearshore stocks.
3. Continue to build evidence for appropriate natural mortality values for females and males. 
4. Improved historical catch reconstructions. Specifically, the historic trawl fishery catches (pre-1987) in particular require particular attention. A synoptic catch reconstruction is recommended, where states work together to resolve cross-boundary state catch issues as well as standardize the approach to catch recommendations to the extent possible.
5. Stock structure for black rockfish is a complicated topic that needs further analysis. How this is determined (e.g., exploitation history, genetics, life history variability, biogeography, etc.) and what this means for management units needs to be further refined. This is a general issue for all nearshore stocks that likely have significant and small scale stock structure among and within states, but limited data collections to support small-scale management.
6. Continue acoustic-visual fisheries independent coastwide survey to develop a time series. Further refine the survey by addressing the recommendations of the SSC methodology review from 2022. Examine the potential of using spatial modeling to reduce the uncertainty in the population estimates from the acoustic-visual fisheries independent coastwide survey. The STAR panel suggested the survey team focus on improving the survey estimates by a) obtaining a target strength estimate for black rockfish, and b) developing a method for in-situ transducer calibration.
7. Reconcile contradictory signal in the Black Rockfish biology versus the population scale.
8. Better understand the ecology and habitats of Black Rockfish from settlement to age 4. Further development of surveys aimed specifically at recruitment or settlement rates of nearshore species, such as OSU’s SMURF collections, that are not frequently encountered in offshore federal age-0 surveys is needed. 
9. Consider additional capacities in Stock Synthesis to model marine reserves (i.e., closed to fishing) and areas that are open to fishing.
10. Simulation analyses or make a standard sensitivity exploration to examine circumstances in which options for treatment sex data for composition data are preferable under Option 1 or 2 treating them as separate or Option 3 treating them as combined and preserving sex ratio within samples.  Such studies should aim to provide criteria for their application to inform guidance in the PFMC’s Groundfish  Terms of  Reference and Accepted Practices documents.
11. Develop and evaluate using the acoustic visual survey data to develop an informative prior for the PIT tag survey.
12. Continue to collect functional maturity information and evaluate the role of geography, environmental forcing, and density dependence on functional maturity estimates for black rockfish. 
