```{r executive, echo = FALSE}
executive <- list()
executive[["stock"]] <- paste0("This assessment reports the status of ",
  spp, " (_", spp.sci, "_) off the ", coast, 
  " coast using data through xxxx.")
```

# Executive summary{-}

## Stock{-}

This assessment reports the status of `r spp` (*Sebastes melanops*) off Oregon state using data through `r endyr`. `r Spp` are also found in California (their core range) and Washington waters of the U.S. West Coast, and those are treated in separate area-based stock assessments given different mangement considerations and exploitation histories as discussed at the preliminary stock assessment workshop in February 2023 [@PFMC_dataworkshop_2023]. There is substantial biogeographic separation in the populations off Oregon and Washington, thus justifying separation of those populations into different management units and stock assessments. `r spp` are also caught from the waters off British Columbia and Alaska, but there have not been any formal assessments of stock status for those areas. Genetic studies of stock structure indicate fish in Alaska are more differentiated than the along the contiguous west coast of the Unites States, and that genetic diversity varies in a nonsystematic way from California to Oregon [@Hessetal_2023].

## Removals{-}

`r spp` are caught by a wide variety of gear types in Oregon and in recent decades have been a very important target species for recreational charter-boats and private sport anglers (i.e., ocean boat fleet).  In recent years the recreational fishery has accounted for most of the `r spp` catches (Figure \ref{fig:es-catch}).  `r spp` can also be an important component of nearshore commercial fisheries, either as incidental catch by the troll fishery for salmon or as directed catch by jig fisheries for groundfish.  Further, there are nearshore commercial fisheries that catch and sell fish live for the restaurant trade.  There has been almost no trawl-caught landings of `r spp` in recent years (Table \ref{tab:removalsES}), but trawl landings in the 1940s to 1970s were more prominant (Figure \ref{fig:es-catch}).

Detailed reports of commercial landings of `r spp` are generally unavailable prior to 1981, when the Pacific Fishery Information Network (PacFIN) database began.  The catch series prior to 1981 for these assessments were derived by applying available estimates or assumed values for the proportion of `r spp` landings in reported landings of rockfish.  Observer data, which are available since the early 2000s, indicate low levels of discarding of `r spp`, generally less than 2% of total catch. While `r spp` are unlikely to have ever comprised a large percentage of overall rockfish landings due to their low abundance compared to other rockfish species, it seems plausible that they have been more than a trivial component due to their nearshore distribution for many years.  `r spp` were one of only four rockfish species mentioned by scientific name in reports of rockfish landings in Oregon during the 1940s.

Overall, removals of `r spp` remained relatively low (less than 100 mt) until the mid to late 1970s when landings quickly quadrupled with the expansion of the recreational fishery. Since the 1980s, removals have consistently fluctuated between 300 and 600 mt (no major trend), comprising mostly of removals from the ocean boat recreational fleet and the non-trawl commercial fleet (Figure \ref{fig:es-catch}).  


\input{tex_tables/a_Catches_ES.tex}


```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "catch2 landings stacked.png"), 
caption = "Landings by fleet used in the reference model where catches in metric tons by fleet are stacked",
label = 'es-catch')
```

\clearpage

## Data and assessment{-}

The first `r spp` stock assessment along the west coast of the United States that included the majority of Oregon waters was completed in 2007, covering the area south of Cape Falcon, Oregon to north of Point Piedros Blancos, California [@Sampson_2007]. In 2015, a subsequent assessment was completed that included Oregon waters only as one of three (also Washington and California) seperate assessment areas delineated by state lines [@Copeetal_2015]. Similarly, this assessment treats Oregon waters as a single assessment area. All of these assessments have used Stock Synthesis software. This assessment used Stock Synthesis version 3.30.21.00.

This assessment integrates data and information from multiple sources into one modeling framework. The stock assessment model for `r spp` is informed by catch data from two commercial fleets and two recreational fleets, six abundance indices, five sets of length composition data, and 3 sets of conditional age-at-length compositions. It also uses an ageing error matrix to incorporate ageing imprecision and applies fixed parameterizations of weight-at-length, maturity-at-length, and fecundity-at-length, the Beverton-Holt stock-recruitment steepness value and recruitment variability. Life history parameters were sex-specific (i.e., a two-sex model) with natural mortality fixed at external estimates and growth and recruitment parameters estimated.  Additional parameters that were estimated include initial population scale ($lnR_0$), selectivity for each fishery and survey, and extra survey variance. The base model was tuned to account for the weighting of the length and age data and index variances (which was estimated), as well as the specification of the recruitment bias adjustments.  Derived quantities include, among other things, the time series of spawning biomass, age and size structure, and current and projected future stock status. The model covers the years `r startyr` to `r endyr`, with a 12 year forecast beginning in `r endyr + 1`.

Within model uncertainty is explicitly included in this assessment by parameter estimation uncertainty, while among model uncertainty is explored through sensitivity analyses addressing alternative input assumptions such as data treatment and weighting, and model specification sensitivity to the treatment of life history parameters, selectivity, recruitment, and survey catchability. A reference model was selected that best fit the observed data while concomitantly balancing the desire to capture the central tendency across those sources of uncertainty, ensure model realism and tractability, and promote robustness to potential model misspecification.

## Stock biomass and dynamics{-}

Spawning output (in millions of eggs; meggs) instead of spawning biomass is used to report the functionally mature population scale because fecundity is nonlinearly related to body female weight. The estimated spawning output at the beginning of 2023 was `r ssb_curr` meggs (~95 percent asymptotic intervals: `r ssb_curr_low` to `r ssb_curr_hi` meggs, Table \ref{tab:ssbES} and Figure \ref{fig:es-ssb}), which when compared to unfished spawning output (`r ssb0`) meggs gives a relative stock status level of `r round(100*depl_curr,0)` percent (~95 percent asymptotic intervals: `r round(100*depl_curr_low,0)` to `r round(100*depl_curr_hi,0)` percent, Figure \ref{fig:es-depl}).  Overall, spawning output declined with the onset of increasing commercial removals in the 1960s and continued to decline with the increase in recreational catches through the 1990s, even dropping below the target relative stock size from 1993 to 2008, before steadily increasing back above target since that time. The largest of the estimated recruitment pulses occurred in 2008 and was followed by several above average recruitment years in the early 2010s, which contributed to the increase in spawning output through the mid to late 2010s. The minimum relative stock size of `r round(100*as.numeric(depl_min),0)` percent of unfished levels is estimated to have occurred in `r depl_min_yr`. Accordingly, the stock has not been below the minimum stock size threshold (i.e., never overfished based on median estimates). Currently the stock is estimated above the management target of $SB_{40\%}$ in 2023 and is estimated to have remained above the target since 2009 (Table \ref{tab:ssbES} and Figure \ref{fig:es-depl}).

\input{tex_tables/b_SSB_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots",
"ts7_Spawning_output_with_95_asymptotic_intervals_intervals.png"), 
caption = "Estimated time series of spawning output (circles and line: median; light broken lines: 95 percent intervals) for the base model",
label = 'es-ssb')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "ts9_Relative_spawning_output_intervals.png"), 
caption = "Estimated time series of fraction of unfished spawning output (circles and line: median; light broken lines: 95 percent intervals) for the base model",
label = 'es-depl')
```

\clearpage

## Recruitment{-}

Recruitment is informed by the data from 1980 to 2017 (Table \ref{tab:recrES} and Figure \ref{fig:es-rec-ts}).  The highest recruitment years occurred in 1999, 2000, 2008, 2013, and 2016. The large 2008 and 2016 year classes, as well as several above average year classes in the early 2010s, contributed to the recent increase in `r spp` biomass. Recruitment is informed by composition data and six relative abundance indices.  The 2015 stock assessment did not estimate deviations from the stock-recruitment curve. While the `r spp` stock has been reduced to levels that theoretically would provide some information on how recruitment compensation changes across spawning biomass levels (i.e., inform the steepness parameter), the assessment model could not adequately estimate a reasonable steepness parameter.  Thus, recruitment is based on a fixed assumption about steepness ($h$ = 0.72) and recruitment variability ($\sigma_R$ = 0.6).

\input{tex_tables/c_Recr_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "ts11_Age-0_recruits_(1000s)_with_95_asymptotic_intervals.png"), 
caption = "Estimated time series of age-0 recruits (1000s) for the base model with 95 percent intervals",
label = 'es-recruits')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "ts11_Age-0_recruits_(1000s)_with_95_asymptotic_intervals.png"), 
caption = "Estimated time series of recruitment deviations",
label = 'es-rec-ts')
```

## Exploitation status{-}

California and Washington models indicate that current fishing practices are near or above the SPR rate fishing intensity target, while the Oregon model is quite a bit above the target (table ES-4, compare to SPR=0.5; Figure ES-13 to Figure ES-18), though the steepness value (0.773) indicates a much lower value of SPR for sustainable removals. Fishing rates have been above the target in California in nearly all years since the 1980s, but have dropped considerably in recent years. Oregon fishing rates have been consistently high in recent years.  Washington shows a dramatic decline in fishing intensity since the late 1990s and has fluctuated mostly below the target since. 

Trends in fishing intensity (1 - SPR) largely mirrored that of landings until the 1990s when recruitment pulses overcame the catches to lower overall fishing intensity (Figure \ref{fig:es-1-spr}). The maximum fishing intensity was `r spr_max` in `r spr_max_yr`, above the target SPR-based harvest rate of 0.50 (1 - $\text{SPR}_{50\%}$). Current levels of `r spr_recent_10[10]` for 2020 are near the fishing limit. Fishing intensity over the past decade has ranged between `r min(spr_recent_10)` and `r max(spr_recent_10)` and the exploitation rate has been high (`r min(exploit_recent_10)` - `r max(exploit_recent_10)`, Table \ref{tab:exploitES}). Current estimates indicate that `r spp` spawning output is much greater than than the target biomass level ($\text{SB}_{40\%}$), though fishing intensity remains near target $F_{MSY}$ proxy harvest rate.

\input{tex_tables/d_SPR_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "SPR2_minusSPRseries.png"), 
caption = "Estimated 1 - relative spawning ratio (SPR) by year for the base model. The management target is plotted as a red horizontal line and values above this reflect harvest in excess of the proxy harvest rate",
label = 'es-1-spr')
```

\clearpage


## Ecosystem considerations{-}

This stock assessment does not explicitly incorporate trophic interactions, habitat factors or environmental factors into the assessment model. More predation, diet and habitat work, and mechanistic linkages to environmental conditions would be needed to incorporate these elements into the stock assessment. While no mechanisms have been put forth for these time-varying changes in growth, an environmental component is possible. Limited data in Oregon and California also suggest the possibility that growth has changed over time.

## Reference points{-}

Reference points were based on the rockfish FMSY proxy (SPR50%), target relative biomass (40%) and model-estimated selectivity for each fleet. California is below the target biomass reference point, but above the limit reference biomass (25%). Oregon is well above the target biomass.  Washington relative biomass is above the target biomass. California and Washington yield values are lower than the previous assessment for similar reference points due to lower overall natural mortality values (Table ES-5). The proxy MSY values of management quantities are the most conservative compared to the estimated MSY and MSY relative to 40% biomass for both California and Washington (Table ES-5). The equilibrium estimates of yield relative to biomass are provided in Figure ES-19 to Figure ES-21.

The 2021 spawning biomass relative to unfished equilibrium spawning biomass is well above the management target of 40 percent of unfished spawning biomass. The relative biomass and the ratio of the estimated SPR to the management target ($\text{SPR}_{50\%}$) across all model years are shown in Figure \ref{fig:es-phase} where warmer colors (red) represent early years and colder colors (blue) represent recent years.  There have been periods where the stock status has decreased below the target and fishing intensity has been higher than the target fishing intensity based on $\text{SPR}_{50\%}$. Figure \ref{fig:es-yield} shows the equilibrium curve based on a steepness value fixed at `r round(model$parameters[model$parameters$Label == "SR_BH_steep","Value"], 2)` with vertical dashed lines to indicate the estimate of fraction unfished at the start of `r endyr +1` (current) and the estimated management targets calculated based on the relative target biomass (B target), the SPR target, and the maximum sustainable yield (MSY).

Reference points were calculated using the estimated selectivities and catch distributions among fleets in the most recent year of the model, `r endyr` (Table \ref{tab:e_ReferencePoints_ES}). Sustainable total yield, removals, using an $\text{SPR}_{50\%}$ is `r format(model$derived_quants[model$derived_quants$Label == 'Dead_Catch_SPR', 'Value'], digits = 3, scientific = FALSE, big.mark = ',')` mt. The spawning output equivalent to 40 percent of the unfished spawning output ($\text{SO}_{40\%}$) calculated using the SPR target ($\text{SPR}_{50\%}$) was `r format(model$derived_quants[model$derived_quants$Label == 'SSB_SPR', 'Value'], digits = 4, scientific = FALSE, big.mark = ',')` meggs. Recent removals have been close to the point estimate of potential long-term yields calculated using an $\text{SPR}_{50\%}$ reference point and the population size has been relatively decreasing toward the target over the past few years.

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "SPR4_phase.png"), 
caption = "Phase plot of estimated 1-SPR versus fraction unfished for the base model",
label = 'es-phase')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "yield2_yield_curve_with_refpoints.png"), 
caption = "Equilibrium yield curve for the base case model. Values are based on the 2020
fishery selectivities and with steepness fixed at 0.80",
label = 'es-yield')
```

\input{tex_tables/e_ReferencePoints_ES.tex}

\clearpage

## Management performance{-}

Removals have been below the equivalent ABC-ACL since the prior assessment (Table ES-6), but those specified ABCs from the 2007 assessments are higher than those coming from the current assessment models. Removals over the last few years have or may have exceeded the newly estimated ABC-ACL values in some years. The differences in the treatment of natural mortality between the previous and current assessments are the biggest reason for this discrepancy.

Exploitation on `r spp` increased starting around 1960 and reached a high in the early 1990s. Since that time, catch has mostly fluctuated between 5 and 10 mt per year, with some years exceeding 10 mt, particularly in the last 4 years. The last ten years of the vermilion rockfish component acceptable biological catch (ABC) and annual catch limit (ACL) (which are equivalent) of the Minor Shelf Rockfish North Complex has been set, by definition, below the overfishing limit (OFL) (Table \ref{tab:ofl-es}). The `r spp` component OFL for this Complex has been exceeded by the Oregon removals in the most recent 4 years.

#```{r, results = 'asis'}
#yrs = hist = 2011:endyr
#nfleets = 2
#catch = dead = est.dead = input.catch =  0
#for (i in 1:nfleets){
#  name = paste0("retain(B):_",i)
#  input.catch = model$timeseries[model$timeseries$Yr %in% yrs, name]
#  catch = cbind(catch, input.catch)

#  name = paste0("dead(B):_",i)
#  est.dead = model$timeseries[model$timeseries$Yr %in% yrs, name]
#  dead = cbind(dead, est.dead)
#}
#total.catch <- round(apply(catch, 1, sum),1)
#total.dead  <- round(apply(dead, 1, sum), 1)

#man = read.csv("C:/Users/Jason.Cope/Documents/Github/Vermilion rockfish OR WA assessment 2021/OR/write_up/tables/#vermilion_man_for_doc.csv")
#man = man[man$Year %in% yrs, ]
#out = cbind(man,  total.catch, total.dead)

#col_names = c("Year", "OFL", "ABC", "ACL", "Landings", "Est. Total Mortality")

#table_format(x = out,
#             caption = "The OFL, ABC, ACL, landings, and the estimated total mortality in metric tons.",
#             label = "ofl-es",
#             align = 'l',
#             col_names = col_names)
#```

## Unresolved problems and major uncertainties{-}

The most significant uncertainty for all models is the treatment and value of natural mortality and the form of fleet selectivity (e.g., length-based asymptotic vs. age-based dome-shaped selectivity). Data-driven selection between the extreme “kill” (using a ramping of M) or “hide” hypotheses are not currently resolvable. The current California and Washington base models instead use a form of the “kill” hypothesis by not implementing the age-based selectivity (“hide” hypothesis) and estimating female and male natural mortality, thus avoiding a fixing natural mortality as was necessary in the Oregon model. The Oregon model also contained a step in female natural mortality, a specification not used in the California or Washington models. Another important issue is the highly uncertain historical time-series of removals in all states, which needs further consideration. The development of fishery-dependent indices of abundance still requires further attention. Steepness, while fixed, is still highly uncertain for rockfishes and currently is mismatched to the MSY proxy. And while the steepness profile shows low sensitivity in several derived quantities, steepness strongly defines the yield capacity of stocks, and therefore could cause major uncertainty in the recommended management quantities. Stock structure and its relationship to the current political/management boundaries are also not fully understood, both within U.S. jurisdiction and between the U.S. and Canada. While this is a common challenge faced in most west coast stock assessments, further improvement on this topic will likely rely on black rockfish-specific data.

## Scientific uncertainty{-}

The model-estimated uncertainty around the `r endyr+1` spawning biomass was $\sigma$ = `r round(model$Pstar_sigma,2)` and the uncertainty around the OFL was $\sigma$ = `r round(model$OFL_sigm, 2)`. This is likely an underestimate of overall uncertainty because of the necessity to fix some parameters such as steepness, as well as a lack of explicit incorporation of model structural uncertainty.

## Harvest Projections and Decision Table{-}

Black rockfish assessments for California and Washington have a preliminary distinction as category 1 stock assessments, thus harvest projections and decision tables are based on using P*=0.45 and sigma = 0.36, resulting in a multiplier on the OFL of 0.956. The Oregon black rockfish assessment is a category 2 assessment, with a P*=0.45 and sigma = 0.72 with a multiplier of 0.913 applied to the OFL.  These multipliers are also combined with the rockfish MSY proxy of FSPR=50% MSY and the 40-10 harvest control rule to calculate OFLs, ABCs and ACLs. Projections for each state are provided in Table ES-7 to Table ES-9. 

Uncertainty in management quantities for the base model of each state was characterized by exploring various model specifications in a decision table. Initial exploration included natural mortality and steepness values, and uncertainty in historical trawl catches for the WA and CA models. OR explored the scale factor coming from the value of the tagging catchability (Q) parameter, as well as M values. For the CA and WA models, there was very little sensitivity to steepness and trawl catches, but natural mortality produced sensitive results of predicted population scale and status. Discussion with the STAR panel resulted in high and low states of nature +/- 0.03 from the base case natural mortality values for females and males. High and low catch streams (rows) were determined by the forecasts, as described above, for each state of nature. Thus the low catch stream is based on the forecast from the low state of nature. The OR model demonstrated little sensitivity to M, but high sensitivity to the tagging survey Q. High and low states of nature, respectively, were based on a fixed tag of Q = 0.125 and Q estimated by the model. Resultant decision tables are provided in Table ES-10 to Table ES-12.

A ten year (2023-2032) projection of the reference model with removals in 2021 and 2022 provided by the Groundfish Management Team for each fleet under the category 1 (sigma=0.5) time-varying buffer using $P^*$ = 0.45 and 40-10 ABC control rule is provided in Table \ref{tab:project_ES}.

 \input{tex_tables/project_ES.tex}

The decision table (Table \ref{tab:es-dec-tab}) was constructed using female and male natural mortality to define the low and high states of nature. The multi-parameter likelihood profile was used to find the low (Female M = 0.07092; Male M= 0.06525) and high (Female M = 0.08527; Male M = 0.07845) female and male natural mortality values that produce -log likeliehood values  +0.66 units from the reference -log likelihood value. These correspond to the 12.5% and 87.5% quantiles (standard quantiles used in west coast decision tables). The catch rows in the table were based on three proposed catch streams:
  1. P* = 0.45, sigma = 0.5
  2. P* = 0.40, sigma = 0.5
  3. An equilibrium catch based on the $F_{MSY}$ proxy using SPR = 0.5.  

Across all states of natures and catch streams, vermilion rockfish relative stock size never falls below the target relative stock size of 40%. Both P* approaches lower the stock status from the high relative stock size values, while the $F_{MSY}$ proxy does not. The mismatch in the corresponding steepness value ($h=0.6$) that matches MSY at SPR = 0.5 with the steepness value in the stock assessment ($h=0.72$) that correpsonds to an MSY SPR of 0.35 explains why this constant catch will maintain the stock at very high relative stock status levels. 

\clearpage

 \input{tex_tables/decision_table_es.tex}

\clearpage


## Research and data needs{-}

Recommended avenues for research to help improve future black rockfish stock assessments:

1. Further investigation into the movement and behavior of older (> age 10) females to reconcile their absence in fisheries data. If the females are currently inaccessible to fishing gear, can we find where they are?
2. Appropriate natural mortality values for females and males. This will help resolve the extent to which dome-shaped age-based selectivity may be occurring for each.
3. All states need improved historical catch reconstructions. The trawl fishery catches in particular require particular attention. Given the huge historical removals of that fleet in each state, the assessment is very sensitive to the assumed functional form of selectivity. A synoptic catch reconstruction is recommended, where states work together to resolve cross-state catch issues as well as standardize the approach to catch recommendations.
4. Identifying stanzas or periods of uncertainty in the historical catch series will aid in the exploration of catch uncertainty in future assessment sensitivity runs. 
5. The ODFW tagging study off Newport should be continued and expanded to other areas.  To provide better prior information on the spatial distribution of the black rockfish stock, further work should be conducted to map the extent of black rockfish habitat and the densities of black rockfish residing there.
6. An independent nearshore survey should be supported in all states to avoid the reliance on fishery-based CPUE indices.
7. Stock structure for black rockfish is a complicated topic that needs further analysis. How this is determined (e.g., exploitation history, genetics, life history variability, biogeography, etc.) and what this means for management units needs to be further refined. This is a general issue for all nearshore stocks that likely have significant and small scale stock structure among and within states, but limited data collections to support small-scale management.

