# Stock Structure
The stock structure of black rockfish continues to be based on state-specific boundaries. Stocks definitions are based on a variety of factors that include genetics, biology, ecology, habitat availability and fishing mortality history. Black rockfish range from southern California up to Alaska. There ecology as a nearshore (i.e., commonly an area of restricted gene flow) rockfish with semi-pelagic behavior (e.g., can demonstrate rare, but substantial adult movements) under strong and differential fishing pressure by state provides an interesting assortment of considerations. A recent genetics study done by Hess et al. (in review) found that Alaskan and west coast populations showed significant genetic difference. The population from Washington to California showed a variety of interesting genetic signatures, with decreased genetic diversity north of Cape Mendocino and North of the Columbia River, but increasing in waters off of Oregon. There were also pockets of isolation by distance and odd signatures off Brookings, Oregon. These variable genetic signatures combined with the lack of consistent black rockfish habitat in southern Washington, and the different exploitation histories in each of the states made state-level designations to support management decisions the most logical. This does not suggest there is no exchange of individuals among California, Oregon and Washington, but instead acknowledges that the exchange is likely low enough not to homogenize the populations, and that different population trends can be expected in each state. This was supported in the results of the last assessment, where important differences in stock status were observed.  

# Fleet Structure

Currently, the following fleet structure is being considered for modeling commercial and recreational fisheries in both area models:
 
1. Commercial fixed-gear Fleet (dead and live fisheries combined)
2. Commercial trawl (mostly a historical fishery)
3. Recreational boat-based 
4. Recreational shore-based

Defining fleets is largely based on whether a fishing approach differs in the selectivity (i.e., the capture of fish by length and/or age). Selectivity translates how the removals are taken via length and/or age out of the population. The above four fishery types are distinguished by different fishing activities that result in different selectivities. While in the past the commercial fixed-gear fisheries had been separated by the condition of the landed fish (dead and live), it has been determined that the designation of how the fish will be delivered is a processing issue, not a capture/selectivity issue (Figure \ref{fig:comm_dead_live}). Thus the gear selectivity is considered the same in these gears. This similarity in selectivity is also apparent when looking at the 2015 assessment selectivity estimates for each fishery. The fixed-gear and trawl fisheries do present different selectivities (Figure \ref{fig:com-length-dist}) justifying different fleets. 

For the recreational fishery, the private and charter boat-based fisheries show similar aggregate length compositions and are treated as one fleet (Figure \ref{fig:private_charter_lts}). The boat- and shore-based fleets show distinct aggregate compositions that demonstrate the well known characteristic of shore-based landings catching smaller fish (Figure \ref{fig:boat_shore_lts}). These recreational fleets are the same treatments as in the 2015 assessment.


```{r, results = 'asis'}
add_figure(
  filein = file.path(dir, "plots", "black_density plot length by gear and condition_fixed.jpg"), 
  caption = "Commerical fixed-gear samples and aggregate length distribution by condition (A = Live; R= Dead).",
  label = 'comm_dead_live')

``` 


```{r, results = 'asis'}
add_figure(
filein = file.path(dir, "plots", "black_density plot length by fleet_with means.png"), 
caption = "Aggregate length compositions by the two main commercial gears.",
label = 'com-length-dist')

```
\


```{r, results = 'asis'}
add_figure(
  filein = file.path(dir, "plots", "Boat_comps.jpg"), 
  caption = "Aggregate length compositions of the private boat and charter boat fisheries. The compositions are not notably different, thus a combined boat-based fishery is used in the assessment.",
  label = 'private_charter_lts')

``` 

```{r, results = 'asis'}
add_figure(
  filein = file.path(dir, "plots", "Boat_Shore.jpg"), 
  caption = "Aggregate length compositions of the boat and shore-based fisheries. The compositions are notably different, thus a boat-based and shore-based fisheries are treated as separate fleets in the assessment.",
  label = 'boat_shore_lts')

``` 
