

# Composition Data

## Commercial length compositions

Samples sizes reflect the fact that the fixed-gear fishery is more prominent than the trawl fishery. The trawl fishery is an amalgam of several contemporary gear types that do not necessarily have similar selectivities, but the removals are so small that parsimony argues for a "trawl" gear of other commercial gears taking black rockfish.


```{r, results = 'asis'}
add_figure(
filein = file.path(dir, "plots", "black commercial_histogram length by fleet.tiff"), 
caption = "Sample sizes of the two commercial fleets.",
label = 'com-length-n')

```
\

```{r, results = 'asis'}
add_figure(
filein = file.path(dir, "plots", "black_density plot length by gear_trawl.jpg"), 
caption = "Aggregate length compositions by gear that comprise the 'trawl' fishery.",
label = 'lt _gear_trawl')

```

\

## Recreational length compositions

The recreational length composition data summarized below represen data pulled from RecFIN collected by either the MRFSS (1980 - 2003) or CRFS (2004 - 2022) sampling programs. There are additional data sources that contain historical length samples from the CPFV fleets (1975-1979 from Collins and Crooke, 1987-1998 from Deb Wilson-Vandenberg, and 1986-1989 from Alley and Ono) that will be evaluated and used within each assessment model as appropriate but are not included here.

